Testing the Overconfidence Trap

### 2023-09-29 Slides/overconfident_beliefs/TheoryLunch_fall22.tex
slides setup

\usepackage{geometry}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{setspace}
\usepackage{theorem}
\usepackage{natbib}
\usepackage{mathtools}
\usepackage{cite}
\usepackage{natbib}
\usepackage{setspace}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{xcolor}
\usepackage{array}
\usepackage{caption}
\usepackage{graphicx}
\usepackage{siunitx}
\usepackage[normalem]{ulem}
\usepackage{colortbl}
\usepackage{multirow}
\usepackage{hhline}
\usepackage{calc}
\usepackage{tabularx}
\usepackage{threeparttable}
\usepackage{wrapfig}
\usepackage{adjustbox}
\usepackage{hyperref}
\usepackage{tikz}
\usepackage[table]{xcolor}

\title{Models of Belief Updating:\\ The Case of Overconfidence}
\author{Jimena Galindo}

@@

\begin{frame}{Motivation}
    We often observe beliefs and behaviors that are inconsistent with the fully Bayesian benchmark:\\
    \bigskip
    \begin{itemize} 
        \item Overconfidence  
        \item Polarization 
    \end{itemize}

    \bigskip
    Recent theoretical literature has sought to explain these phenomena (and others) with the notion of \bf{Misspecified Mental Models}
\end{frame}

\begin{frame}{Motivation}
    We often observe beliefs and behaviors that are inconsistent with the fully Bayesian benchmark:\\
    \bigskip
    \begin{itemize} 
        \item Overconfidence \textbf{(today)}  
        \item Polarization \textbf{(next time)}
    \end{itemize}

    \bigskip
    Recent theoretical literature has sought to explain these phenomena (and others) with the notion of \bf{Misspecified Mental Models}
\end{frame}

\begin{frame}{Mental Models}

A prior on unknown parameters/states and an updating procedure\\
\bigskip
\begin{itemize}
    \item Bayesian 
    \item Dogmatic 
    \item Motivated Beliefs
\end{itemize}

@@

\begin{frame}{Overconfidence}
    Belief that my type is higher than it truly is (``overestimation'' as in Moore and Healy, 2008)\\
    \bigskip

    Overconfidence seems to be persistent in various settings. Ultimately it leads to sub-optimal choices:
    \begin{itemize} 
        \item Excess entry of entrepreneurs (Camerer and Lovallo, 1999)
        \item Suboptimal genetic testing (Oster et al. 2013)
        \item  Workers overestimate their productivity (Hoffman and Burks, 2020)
    \end{itemize}

\begin{frame}{An Example}
    An entrepreneur has \textbf{unknown intrinsic ability} $\theta^*$ and chooses a level of effort $e\geq 0.$ \\
    \bigskip
    Effort and ability are transformed into output at an exogenous and \textbf{unknown rate} $\omega.$\\
    \bigskip 
    An overconfident entrepreneur believes he is of type $\hat\theta>\theta^*$\\
    \bigskip

    The entrepreneur wants to maximize their utility\\
        $$y = (\theta^* + e)\omega-\frac{1}{2}e^2 +\varepsilon$$\\
    \pause
    \bigskip
    Regardless of their own type (or their beliefs about it), they should choose $e^*(\omega)=\omega$\\
\end{frame}

\begin{frame}{Learning is Possible}
    This exercise is repeated for $t=0, 1, ...$
        $$y_t = (\theta^* + e_t)\omega-\frac{1}{2}e_t^2 +\varepsilon_t$$

    Note that both parameters are identified in this setting:\\
    \bigskip

    \begin{itemize}
        \item Choosing $\hat{e}$ and $\hat{e}+1$ over multiple periods allows identification of $\omega$\\
        \bigskip
        \item Once $\omega$ is known, $\theta$ can be backed out\\
     \end{itemize}
    \bigskip
    How come people don't learn the true $\theta$?
\end{frame}

\begin{frame}{The Literature}
Settings with two or more unknowns allow for different explanations of the bias:\\
\bigskip
\begin{enumerate}
\item Self-defeating equilibrium (Heidhues et al., 2018): 
    \begin{itemize}
        \item Bayesian on $\omega$
        \item Dogmatic about $\theta$

    \end{itemize}
    \bigskip
    \item Bayesian Likelihood Ratio test (Ba, 2022 JMP):
    \begin{itemize}
        \item Bayesian on $\omega$ 
        \item Hypothesis testing on $\theta$
    \end{itemize}
    \bigskip
    \item Self-Serving Attribution Bias with two unknowns (Brunnermeier and Parker, 2005; Coutts et al. 2022wp): 
    \begin{itemize}
        \item Good news are attributed to high $\theta$ bad news are attributed to low $\omega$
    \end{itemize}

@@

\begin{frame}{The Research Questions}
    We know people fall somewhere between Bayesian Learning and the very extreme Heidhues et al. setting.\\
    \bigskip
    \begin{itemize}
        \item To what extent do the different theories explain the data?\\
        \begin{itemize}
            \item Is there heterogeneity in the way people update beliefs?\\
        \end{itemize}
    \end{itemize}
    \pause
    \bigskip
    Misspecified mental models lead to incorrect beliefs even in settings where there are no ego-relevant parameters
    \bigskip
    \begin{itemize}
        \item How do the models perform in a setting with two parameters that are NOT ego-relevant?\\
    \end{itemize}
\end{frame}

\begin{frame}{Roadmap}
    \begin{enumerate}
        \item Theory 1 --- Further Motivation\\
        \bigskip
        \item Theory 2\\
        \bigskip
        \item Theory 3\\
        \bigskip
        \item Unifying framework\\
        \bigskip
        \item Experimental Design
    \end{enumerate}
\end{frame}

\begin{frame}{Theory 1: Dogmatic Modelers (HKS)}
    \Large\textbf{Unrealistic Expectations and Misguided Learning \\}
    (Heidhues, Köszegi, and Strack, 2018)
\end{frame}

@@

$\omega$ is drawn from density $g_0$ with $\omega^*=E_{g_0}(\omega).$\\
\bigskip
The entrepreneur's true ability is $\theta^*$, they believe with certainty that it is $\hat\theta>\theta^*.$ \\
\bigskip
At $t=0$, the entrepreneur has the prior $g_0.$\\
\bigskip
They correctly choose $e_0 = \omega^*.$\\
\bigskip
\pause
Suppose they don't update their beliefs or their choice for a long time.
\end{frame}

@@

For their chosen effort $\omega^*$, they observe an average output of 
$$ y_0=(\theta^* + \omega^*)\omega^*-\frac{1}{2}(\omega^*)^2 $$

@@

\begin{frame}{In the Lab}
\label{Figure1}

@@

\begin{figure}
    \centering
    \includegraphics[scale=0.5]{figures/GK_experiment.png}
\end{figure}
    \hyperlink{Other Figures}{\beamerbutton{Other Figures}}

@@

\begin{frame}{Results}
    Biased updating in the same direction as the theoretical predictions.\\
    \bigskip
    The bias is not as large as the theory predicts.\\
    \bigskip
    Some agents do update their belief of $\theta$ in the right direction.\\
\end{frame}

\begin{frame}{Theory 2: Switchers}
    \Large\textbf{ Robust Misspecified Models and Paradigm Shifts \\}
    (Ba, 2022 JMP)
\end{frame}

\begin{frame}{The Setting}
    Same as HKS but with finite $\Omega$ and finite $A$\\
    \bigskip
    Now the entrepreneur is willing to switch to an alternative level of ability $\theta'$ (assume $\theta' = \theta^*$).\\
    \bigskip
    Instead of updating $P[\theta]$ every period, they perform a Bayesian hypothesis test:\\
    \bigskip
    Adopt model $\theta'$ at time t iff \\
    $$\frac{\ell_t (\theta')}{\ell_t(\theta)}>\alpha\geq1$$

@@

    The self-defeating trap is difficult to escape when the entrepreneur is overconfident.\\
    \bigskip
    Underconfident agents escape the trap at a higher rate than overconfident and learn the true parameter.\\
    \bigskip
    Explains why underconfidence is less prevalent in the data.\\
    \bigskip
    Overconfident agents can escape the trap if their prior is not too ``tight'' around a self confirming equilibrium.

@@

\begin{frame}{The Setting}
    A fixed effort $e$, $\theta \in \{\theta_H, \theta_L\}$ and $\omega \in \{\omega_H, \omega_L\}$ generate binary signals (success/failure)\\
    \bigskip
    After a success the agent updates their belief about $\theta$ with distortions $\gamma_\theta$ and $\gamma_\omega$, $p_{t+1}(\theta_H|success)$ is given by:\\
    \begin{center}
        \small
        $\frac{\gamma^s_\theta\gamma^s_\omega\pi_t(\theta_H) g_t(\omega_H) q_t(\theta_H, \omega_H) + \gamma^s_\theta\pi_t(\theta_H) g_t(\omega_L) q_t(\theta_H, \omega_L)}{\gamma^s_\theta\gamma^s_\omega\pi_t(\theta_H) g_t(\omega_H) q_t(\theta_H, \omega_H) + \gamma^s_\theta\pi_t(\theta_H) g_t(\omega_L) q_t(\theta_H, \omega_L) + \gamma^s_\omega\pi_t(\theta_L) g_t(\omega_H) q_t(\theta_L, \omega_H) + \pi_t(\theta_L) g_t(\omega_L) q_t(\theta_L, \omega_L)}$\\
    \end{center}
    \bigskip
    with $q_t(\theta, \omega):=Pr[\text{success}|\theta, \omega]$ increasing in $\omega$ and $\theta$
\end{frame}

\begin{frame}{Predictions and Evidence (Coutts et al. 2022)}
    Prediction: Even unbiased agents will overweight $\theta_H$ after a success and end up being biased.\\
    \bigskip
    Experiment: Evidence of biased updating when $\theta$ is ego-relevant.\\
    \bigskip
    The framework does not allow direct comparisons with the other two theoretical predictions.\\
    \bigskip
    It gives an alternative rationalization for the persistence of overconfidence.

@@

\begin{frame}{A framework where we can compare}

Finite type space: $\theta \in \{\theta_H, \theta_M, \theta_L\}$\\
\bigskip
Finite state space: $\omega \in \{\omega_H, \omega_M, \omega_L\}$
with $p(\omega_k)=1/3$ \\
\bigskip
Finite action space: $e \in \{e_H, e_M, e_L\}$\\
\bigskip
Binary signal: Success/Failure with $P\left[Success|e, \omega, \theta\right]$ satisfying the assumptions of HKS

@@

\begin{frame}{The Data Generating Process}
The probability of success is given by:\\
\bigskip
\centering
\begin{tabular}{ c|c|c|c|}

  \multicolumn{1}{c}{} & \multicolumn{1}{c}{$\omega_H$} & \multicolumn{1}{c}{$\omega_M$} & \multicolumn{1}{c}{$\omega_L$}\\
  \cline{2-4}
  $e_H$ & 50 & 20 & 2 \\
  \cline{2-4}
  $e_M$ & 45 & 30 & 7 \\
  \cline{2-4}
  $e_L$ & 40 & 25 & 20 \\
  \cline{2-4}
  \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{$\theta_L$} & \multicolumn{1}{c}{}\\
\end{tabular}
\hspace{.3cm} % adjust this value to set the space between tables
\begin{tabular}{ c|c|c|c|}

  \multicolumn{1}{c}{} & \multicolumn{1}{c}{$\omega_H$} & \multicolumn{1}{c}{$\omega_M$} & \multicolumn{1}{c}{$\omega_L$}\\
  \cline{2-4}
  $e_H$ & 80 & 50 & 5 \\
  \cline{2-4}
  $e_M$ & 69 & 65 & 30 \\
  \cline{2-4}
  $e_L$ & 65 & 45 & 40 \\
  \cline{2-4}
  \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{$\theta_M$} & \multicolumn{1}{c}{}\\
\end{tabular}
\hspace{.3cm} % adjust this value to set the space between tables
\begin{tabular}{ c|c|c|c|}

  \multicolumn{1}{c}{} & \multicolumn{1}{c}{$\omega_H$} & \multicolumn{1}{c}{$\omega_M$} & \multicolumn{1}{c}{$\omega_L$}\\
  \cline{2-4}
  $e_H$ & 98 & 65 & 25 \\
  \cline{2-4}
  $e_M$ & 80 & 69 & 35 \\
  \cline{2-4}
  $e_L$ & 75 & 55 & 45 \\
  \cline{2-4}
  \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{$\theta_H$} & \multicolumn{1}{c}{}\\
\end{tabular}

@@

\begin{frame}{The Data Generating Process}
    \centering
\begin{tabular}{ c|c|c|c|}

  \multicolumn{1}{c}{} & \multicolumn{1}{c}{$\omega_H$} & \multicolumn{1}{c}{$\omega_M$} & \multicolumn{1}{c}{$\omega_L$}\\
  \cline{2-4}
  $e_H$ & \cellcolor{blue!25}50 & 20 & 2 \\
  \cline{2-4}
  $e_M$ & 45 & \cellcolor{blue!25}30 & 7 \\
  \cline{2-4}
  $e_L$ & 40 & 25 & \cellcolor{blue!25}20 \\
  \cline{2-4}
  \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{$\theta_L$} & \multicolumn{1}{c}{}\\
\end{tabular}
\hspace{.3cm} % adjust this value to set the space between tables
\begin{tabular}{ c|c|c|c|}

  \multicolumn{1}{c}{} & \multicolumn{1}{c}{$\omega_H$} & \multicolumn{1}{c}{$\omega_M$} & \multicolumn{1}{c}{$\omega_L$}\\
  \cline{2-4}
  $e_H$ & \cellcolor{blue!25}80 & 50 & 5 \\
  \cline{2-4}
  $e_M$ & 69 & \cellcolor{blue!25}65 & 30 \\
  \cline{2-4}
  $e_L$ & 65 & 45 & \cellcolor{blue!25}40 \\
  \cline{2-4}
  \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{$\theta_M$} & \multicolumn{1}{c}{}\\
\end{tabular}
\hspace{.3cm} % adjust this value to set the space between tables
\begin{tabular}{ c|c|c|c|}

  \multicolumn{1}{c}{} & \multicolumn{1}{c}{$\omega_H$} & \multicolumn{1}{c}{$\omega_M$} & \multicolumn{1}{c}{$\omega_L$}\\
  \cline{2-4}
  $e_H$ & \cellcolor{blue!25}98 & 65 & 25 \\
  \cline{2-4}
  $e_M$ & 80 & \cellcolor{blue!25}69 & 35 \\
  \cline{2-4}
  $e_L$ & 75 & 55 & \cellcolor{blue!25}45 \\
  \cline{2-4}
  \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{$\theta_H$} & \multicolumn{1}{c}{}\\
\end{tabular}
\end{frame}

\begin{frame}{The Data Generating Process}
\begin{center}
\begin{tikzpicture}
  \draw[<-] (1,0) -- (3,0);
  \draw[<-] (5,0) -- (7,0);
  \draw[<-] (9,0) -- (11,0);
\end{tikzpicture}
\end{center}
\\
\bigskip
\centering
\begin{tabular}{ c|c|c|c|}

  \multicolumn{1}{c}{} & \multicolumn{1}{c}{$\omega_H$} & \multicolumn{1}{c}{$\omega_M$} & \multicolumn{1}{c}{$\omega_L$}\\
  \cline{2-4}
  $e_H$ & \cellcolor{blue!25}50 & 20 & 2 \\
  \cline{2-4}
  $e_M$ & 45 & \cellcolor{blue!25}30 & 7 \\
  \cline{2-4}
  $e_L$ & 40 & 25 & \cellcolor{blue!25}20 \\
  \cline{2-4}
  \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{$\theta_L$} & \multicolumn{1}{c}{}\\
\end{tabular}
\hspace{.3cm} % adjust this value to set the space between tables
\begin{tabular}{ c|c|c|c|}

  \multicolumn{1}{c}{} & \multicolumn{1}{c}{$\omega_H$} & \multicolumn{1}{c}{$\omega_M$} & \multicolumn{1}{c}{$\omega_L$}\\
  \cline{2-4}
  $e_H$ & \cellcolor{blue!25}80 & 50 & 5 \\
  \cline{2-4}
  $e_M$ & 69 & \cellcolor{blue!25}65 & 30 \\
  \cline{2-4}
  $e_L$ & 65 & 45 & \cellcolor{blue!25}40 \\
  \cline{2-4}
  \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{$\theta_M$} & \multicolumn{1}{c}{}\\
\end{tabular}
\hspace{.3cm} % adjust this value to set the space between tables
\begin{tabular}{ c|c|c|c|}

  \multicolumn{1}{c}{} & \multicolumn{1}{c}{$\omega_H$} & \multicolumn{1}{c}{$\omega_M$} & \multicolumn{1}{c}{$\omega_L$}\\
  \cline{2-4}
  $e_H$ & \cellcolor{blue!25}98 & 65 & 25 \\
  \cline{2-4}
  $e_M$ & 80 & \cellcolor{blue!25}69 & 35 \\
  \cline{2-4}
  $e_L$ & 75 & 55 & \cellcolor{blue!25}45 \\
  \cline{2-4}
  \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{$\theta_H$} & \multicolumn{1}{c}{}\\
\end{tabular}

\begin{center}
    \resizebox{0.6\linewidth}{!}{\vector(1,0){300}}
  \end{center}

\begin{frame}{A Self-Confirming Equilibrium}

\centering
\begin{tabular}{ c|c|c|c|}

  \multicolumn{1}{c}{} & \multicolumn{1}{c}{$\omega_H$} & \multicolumn{1}{c}{$\omega_M$} & \multicolumn{1}{c}{$\omega_L$}\\
  \cline{2-4}
  $e_H$ & \cellcolor[HTML]{b84f79}50 & 20 & 2 \\
  \cline{2-4}
  $e_M$ & 45 & 30 & 7 \\
  \cline{2-4}
  $e_L$ & 40 & 25 & 20 \\
  \cline{2-4}
  \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{$\theta_L$} & \multicolumn{1}{c}{}\\
\end{tabular}
\hspace{.3cm} % adjust this value to set the space between tables
\begin{tabular}{ c|c|c|c|}

  \multicolumn{1}{c}{} & \multicolumn{1}{c}{$\omega_H$} & \multicolumn{1}{c}{$\omega_M$} & \multicolumn{1}{c}{$\omega_L$}\\
  \cline{2-4}
  $e_H$ & 80 & \cellcolor[HTML]{b84f79}50 & 5 \\
  \cline{2-4}
  $e_M$ & 69 &\cellcolor[HTML]{f09ebe}65 & 30 \\
  \cline{2-4}
  $e_L$ & 65 & \cellcolor[HTML]{f09ebe}45 & 40 \\
  \cline{2-4}
  \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{$\theta_M$} & \multicolumn{1}{c}{}\\
\end{tabular}
\hspace{.3cm} % adjust this value to set the space between tables
\begin{tabular}{ c|c|c|c|}

  \multicolumn{1}{c}{} & \multicolumn{1}{c}{$\omega_H$} & \multicolumn{1}{c}{$\omega_M$} & \multicolumn{1}{c}{$\omega_L$}\\
  \cline{2-4}
  $e_H$ & 98 & 65 & 25 \\
  \cline{2-4}
  $e_M$ & 80 & 69 & 35 \\
  \cline{2-4}
  $e_L$ & 75 & 55 & 45 \\
  \cline{2-4}
  \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{$\theta_H$} & \multicolumn{1}{c}{}\\
\end{tabular}
\end{frame}

\begin{frame}{The Self-Confirming Equilibria}

\centering
\begin{tabular}{ c|c|c|c|}

  \multicolumn{1}{c}{} & \multicolumn{1}{c}{$\omega_H$} & \multicolumn{1}{c}{$\omega_M$} & \multicolumn{1}{c}{$\omega_L$}\\
  \cline{2-4}
  $e_H$ & \cellcolor[HTML]{b84f79}50 & 20 & 2 \\
  \cline{2-4}
  $e_M$ & 45 & \cellcolor[HTML]{5f94b8}30 & 7 \\
  \cline{2-4}
  $e_L$ & \cellcolor[HTML]{69a35b}40 & 25 & 20 \\
  \cline{2-4}
  \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{$\theta_L$} & \multicolumn{1}{c}{}\\
\end{tabular}
\hspace{.3cm} % adjust this value to set the space between tables
\begin{tabular}{ c|c|c|c|}

  \multicolumn{1}{c}{} & \multicolumn{1}{c}{$\omega_H$} & \multicolumn{1}{c}{$\omega_M$} & \multicolumn{1}{c}{$\omega_L$}\\
  \cline{2-4}
  $e_H$ & 80 & \cellcolor[HTML]{b84f79}50 & 5 \\
  \cline{2-4}
  $e_M$ & \cellcolor[HTML]{fab143}69 & 65 & \cellcolor[HTML]{5f94b8}30 \\
  \cline{2-4}
  $e_L$ & 65 & \cellcolor[HTML]{9662f0}45 & \cellcolor[HTML]{69a35b}40 \\
  \cline{2-4}
  \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{$\theta_M$} & \multicolumn{1}{c}{}\\
\end{tabular}
\hspace{.3cm} % adjust this value to set the space between tables
\begin{tabular}{ c|c|c|c|}

  \multicolumn{1}{c}{} & \multicolumn{1}{c}{$\omega_H$} & \multicolumn{1}{c}{$\omega_M$} & \multicolumn{1}{c}{$\omega_L$}\\
  \cline{2-4}
  $e_H$ & 98 & 65 & 25 \\
  \cline{2-4}
  $e_M$ & 80 & \cellcolor[HTML]{fab143}69 & 35 \\
  \cline{2-4}
  $e_L$ & 75 & 55 & \cellcolor[HTML]{9662f0}45 \\
  \cline{2-4}
  \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{$\theta_H$} & \multicolumn{1}{c}{}\\
\end{tabular}
\end{frame}

\begin{frame}{An Example}
\begin{itemize}
    \item True type is $\theta_M$ \\
    \bigskip
    \item True exchange rate is $\omega_M$ $\rightarrow$ The entrepreneur believes it is uniformly distributed\\
    \end{itemize}

    \centering
\begin{tabular}{ c|c|c|c|}

  \multicolumn{1}{c}{} & \multicolumn{1}{c}{$\omega_H$} & \multicolumn{1}{c}{$\omega_M$} & \multicolumn{1}{c}{$\omega_L$}\\
  \cline{2-4}
  $e_H$ & 50 & 20 & 2 \\
  \cline{2-4}
  $e_M$ & 45 & 30 & 7 \\
  \cline{2-4}
  $e_L$ & 40 & 25 & 20 \\
  \cline{2-4}
  \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{$\theta_L$} & \multicolumn{1}{c}{}\\
\end{tabular}
\hspace{.3cm} % adjust this value to set the space between tables
\begin{tabular}{ c|c|c|c|}

  \multicolumn{1}{c}{} & \multicolumn{1}{c}{$\omega_H$} & \multicolumn{1}{c}{$\omega_M$} & \multicolumn{1}{c}{$\omega_L$}\\
  \cline{2-4}
  $e_H$ & 80 & \cellcolor{blue!25}50 & 5 \\
  \cline{2-4}
  $e_M$ & 69 & \cellcolor{blue!25}65 & 30 \\
  \cline{2-4}
  $e_L$ & 65 & \cellcolor{blue!25}45 & 40 \\
  \cline{2-4}
  \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{$\theta_M$} & \multicolumn{1}{c}{}\\
\end{tabular}
\hspace{.3cm} % adjust this value to set the space between tables
\begin{tabular}{ c|c|c|c|}

  \multicolumn{1}{c}{} & \multicolumn{1}{c}{$\omega_H$} & \multicolumn{1}{c}{$\omega_M$} & \multicolumn{1}{c}{$\omega_L$}\\
  \cline{2-4}
  $e_H$ & 98 & 65 & 25 \\
  \cline{2-4}
  $e_M$ & 80 & 69 & 35 \\
  \cline{2-4}
  $e_L$ & 75 & 55 & 45 \\
  \cline{2-4}
  \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{$\theta_H$} & \multicolumn{1}{c}{}\\
\end{tabular}

>>>>

\usepackage{geometry}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{setspace}
\usepackage{theorem}
\usepackage{natbib}
\usepackage{mathtools}
\usepackage{cite}
\usepackage{natbib}
\usepackage{setspace}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{xcolor}
\usepackage{array}
\usepackage{caption}
\usepackage{graphicx}
\usepackage{siunitx}
\usepackage[normalem]{ulem}
\usepackage{colortbl}
\usepackage{multirow}
\usepackage{hhline}
\usepackage{calc}
\usepackage{tabularx}
\usepackage{threeparttable}
\usepackage{wrapfig}
\usepackage{adjustbox}
\usepackage{hyperref}

\title{Testing the Overconfidence Trap}

@@

\begin{frame}{Motivation}
    Overconfidence seems to be persistent in various settings. Ultimately it leads to suboptimal choices:
    \begin{itemize} 
        \item Excess entry of entrepreneurs (Camerer and Lovallo, 1999)
        \item Suboptimal genetic testing and healthcare (Oster et al. 2013)
        \item  Workers overestimate their productivity (Hoffman and Burks, 2020)
    \end{itemize}
    \bigskip

@@

\begin{frame}{An Example}
    An entrepreneur has \textbf{unknown intrinsic ability} $\theta$ and chooses a level of effort $e\geq 0.$ \\
    \bigskip
    An overconfident entrepreneur believes $\theta$ is larger than the true value.\\
    \bigskip
    Effort and ability are transformed into output at an exogenous and \textbf{unknown rate} $\omega.$\\
    \bigskip 
    So the entrepreneur wants to maximize \\
        $$y = (\theta + e)\omega-\frac{1}{2}e^2 +\varepsilon$$
    \pause
    Regardless of their own type (or their beliefs about it), they should choose $e^*.(\omega)=\omega$\\
\end{frame}

\begin{frame}{Learning is Possible}
    This exercise is myopically repeated for $t=0, 1, ...$
        $$y_t = (\theta + e_t)\omega-\frac{1}{2}e_t^2 +\varepsilon_t$$
    Note that both parameters are identified in this setting:\\
    \bigskip
    \begin{itemize}
        \item Choosing $\hat{e}$ and $\hat{e}+1$ allows identification of $\omega$\\
        \bigskip
        \item Once $\omega$ is known, $\theta$ can be backed out\\
     \end{itemize}
    \bigskip
    How come people don't learn the true $\theta$?
\end{frame}

\begin{frame}{The Literature}
Settings with two or more unknowns allow for different explanations of the bias:\\
\bigskip
\begin{enumerate}
    \item Self-attribution bias with two unknowns (Coutts et al. 2022 wp): 
    \begin{itemize}
        \item Good news are attributed to high $\theta$ bad news are attributed to low $\omega$
    \end{itemize}
    \bigskip
    \item Bayesian factor test (Ba, 2022 JMP):
    \begin{itemize}
        \item Bayesian updating on $\omega$ 
        \item Hypothesis testing on $\theta$
    \end{itemize}
    \bigskip
    \item Self-defeating equilibrium (Heidhues et al., 2018): 
    \begin{itemize}
        \item Never updates beliefs about $\theta$
        \item Bayesian on $\omega$
    \end{itemize}

@@

\begin{frame}{Theory 1}
    \Large\textbf{Unrealistic Expectations and Misguided Learning \\}
    (Heidhues, Köszegi, and Strack, 2018)
\end{frame}

@@

$\omega$ is drawn from density $g_0$. And the realized value is $\omega^*=E_{g_0}(\omega).$\\
\bigskip
The entrepreneur's true ability is $\theta^*$, they believe with certainty that it is $\hat\theta>\theta^*.$ \\
\bigskip
At $t=0$, the entrepreneur has the prior $g_0.$\\
\bigskip
They correctly choose $e_0 = \omega^*.$\\
\bigskip
\pause
Suppose they don't update their beliefs or their choice for a long time.
\end{frame}

@@

They observe an average output of 
$$ y_0=(\theta^* + \omega^*)\omega^*-\frac{1}{2}(\omega^*)^2 $$

@@

\begin{frame}{In the Lab}

@@

\begin{figure}
    \centering
    \includegraphics[scale=0.7]{figures/GK_experiment.png}
\end{figure}

@@

\begin{frame}{Results}
    Biased updating in the same direction as the theoretical predictions.\\
    \bigskip
    The bias is not as large as the theory predicts.\\
    \bigskip
    They argue that some agents do update their belief in the ability parameter.\\
    \bigskip
    How are people updating their beliefs on two parameters?
\end{frame}

\begin{frame}{The Research Question}
    We know people fall somewhere between Bayesian Learning and the very extreme Heidhues et al. setting.\\
    \bigskip
    \begin{itemize}
        \item Do the existing alternative theories explain the behavior well?\\
        \begin{itemize}
            \item Is there heterogeneity in how people update their beliefs? (Bayesian, self-attribution, hypothesis testing, self-defeating)
        \end{itemize}
    \bigskip
        \item Can we design an intervention that will help people learn their true type?\\
    \end{itemize}
\end{frame}

\begin{frame}{Theory 2}
    \Large\textbf{ Robust Misspecified Models and Paradigm Shifts \\}
    (Ba, 2022 JMP)
\end{frame}

\begin{frame}{The Setting}
    Same as Theory 1\\
    \bigskip
    Now the entrepreneur entertains an alternative level of ability $\theta'$ (assume $\theta' = \theta^*$).\\
    \bigskip
    Instead of updating $P[\theta]$ every period, they perform a Bayesian hypothesis test:\\
    \bigskip
    Adopt model $\theta'$ at time t iff \\
    $$\frac{\ell_t (\theta')}{\ell_t(\theta)}>\alpha\geq1$$

@@

    The self-defeating trap is difficult to escape when the entrepreneur is overconfident.\\
    \bigskip
    Underconfident agents escape the trap at a higher rate than overconfident and learn the true parameter.\\
    \bigskip
    Explains why underconfidence is less prevalent in the data.\\
    \bigskip
    Overconfident agents can escape the trap if their prior is not too "tight" around a self confirming equilibrium.

@@

\begin{frame}{The Setting}
    Assume $e$ is fixed and $\theta \in \{\underline{\theta}, \Bar{\theta}\}$ \\
    \bigskip
    Let the probability of success be increasing in $\theta$ and $\omega$ and denote it by by
    $$q(\theta, \omega):=Pr[success|\theta, \omega]$$
    After a success the agent updates their belief about $\theta$ with distortion $\gamma_\theta$:
    $$p_{t+1}(\Bar{\theta}|success)= \frac{\gamma_\theta\pi_t(\Bar{\theta})\sum_\omega g_t(\omega) q_t(\Bar\theta, \omega)}{\gamma_\theta\pi_t(\Bar{\theta})\sum_\omega g_t(\omega)q_t(\Bar{\theta}, \omega)+ \pi_t(\underline {\theta})\sum_\omega g_t(\omega) q_t(\underline{\theta}, \omega)}$$

@@

\begin{frame}{Predictions and Evidence (Coutts et al. 2022)}
    Prediction: Even unbiased agents will overweight $\theta
_H$ after a success and end up being biased.\\
    \bigskip
    Experiment: Evidence of biased updating when $\theta$ is ego-relevant.\\
    \bigskip
    The framework does not allow direct comparisons with the other two theoretical predictions.\\
    \bigskip
    It rationalizes the formation of overconfident beliefs.

@@

\begin{frame}{A framework where we can compare}
    Theories 1 and 2 draw conclusions within the same setting already.\\
    \bigskip
    Can we bring the self-serving attribution bias into that setting?\\
    \begin{enumerate}
        \item A larger $\alpha$ in Theory 2 is consistent with self-serving beliefs when the update is done by evaluating the likelihood ratio.
        \item We can write out the fully Bayesian update within the setting and introduce the bias as in Theory 3.
        \item If some subject starts out with the correct belief $\theta$ and ends up being overconfident, it is evidence of a self-serving bias.
    \end{enumerate}
\end{frame}

\begin{frame}{A Theory-Inspired intervention}
    The parameters in this setting are identified. Allowing for a trade-off between exploitation and experimentation is sufficient for learning. \\
    \bigskip
    There are two ways in which experimentation can lead to correct learning:\\
    \begin{enumerate}
        \item Experimentation on $e$:
        \begin{itemize}
            \item Götte and Kozakiewicz show that some agents do learn (are they experimenting?)
            \item Would overconfident/underconfident subjects choose to experiment if the option is made explicitly available to them?
        \end{itemize}
        \bigskip
        \item Allowing experimentation over $\omega$ (switching environments): 
        \begin{itemize}
            \item Hesterman and Le Yaouanq (2021) show that in a model with fixed $e$ and the possibility to experiment in different environments, overconfident subjects learn $\theta$ faster than unbiased/underconfident subjects
        \end{itemize}
    \end{enumerate}
\end{frame}

\begin{frame}{Testing in the lab}

Part 1: Test\\
\begin{itemize}
    \item Get the ego-relevant parameter
    \item Elicit beliefs
\end{itemize}
\pause
Part 2: Investment game (N rounds)\\
\begin{itemize}
    \item Treatment is paid according to their own score as $\theta$ 
    \item Control is paid according to an ego-neutral parameter 
\end{itemize}
\begin{enumerate}
    \item Choose an amount e to invest
    \item Returns are $\theta \omega + e(\omega-\frac{1}{2}e) + \varepsilon_t$
    \item Belief elicitation
\end{enumerate}
\pause
Part 3: Explicitly allow experimentation
\begin{itemize}
    \item Draw a random $e$ for the next $n$ rounds
    \item Draw a new return parameter $\omega$
\end{itemize}

@@

\begin{frame}{What needs to be done}
    \begin{enumerate}
        \item Define self-serving attribution bias in the framework of theories 1 and 2 (??)\\
        \bigskip
        \item Get predictions in a simplified environment (finite parameter sets).
        \bigskip
        \item Parameters should be such that there are differences among models:
        \begin{itemize}
            \item Subjects update in Theory 2 but fall into self-defeating eq. in Theory 1
            \item Self-serving beliefs make the default model more sticky in Theory 2 
            \item Self serving bias over weights high $\theta$ relative to the Bayesian
            \item Under self serving bias, unbiased subjects become biased. And overconfident subjects become more confident after a success
        \end{itemize}
    \end{enumerate}
\end{frame}
